---
title: "Data Analysis Project"
author: "Amanda Hernandez"
date: "4/16/2022"
output: html_notebook
---

### Loading Packages

```{r}

# look into pop growth in comparison to incarceration rates over time

library(tidyverse)
library(tidycensus)
library(dplyr)
library(sf)
library(tigris)
library(ggplot2)

census_api_key("8b69f2765f65670c2183febdffce6afc73c72101", install=TRUE, overwrite=TRUE)
readRenviron("~/.Renviron")

```

### Data Prep & Exploration

```{r}

incarcerated_pop <- read_csv("data/census_incarceration.csv") %>%
  rename(fips = FIPS)

nrow(incarcerated_pop) # 3211 records
ncol(incarcerated_pop) # 9 columns/variables

# add percent of pop incarcerated (2000, 2010, 2020)

incarcerated_pop <- incarcerated_pop %>%
  mutate(percent_incarcerated_20 = (incarcerated_20/total_population_20)*100) %>%
  mutate(percent_incarcerated_10 = (incarcerated_10/total_population_10)*100) %>%
  mutate(percent_incarcerated_00 = (incarcerated_00/total_population_00)*100)

incarcerated_pop %>%
  arrange(desc(percent_incarcerated_20)) %>%
  select(state, county, percent_incarcerated_20)

# 2020
# CO	Crowley County	48.1256332		
# PA	Forest County	38.0467518		
# MS	Issaquena County	36.8460389		
# LA	West Feliciana Parish	34.5133899		
# GA	Wheeler County 31.6155802

mean(incarcerated_pop$percent_incarcerated_20) # = 1.404581

incarcerated_pop %>%
  arrange(desc(percent_incarcerated_10)) %>%
  select(state, county, percent_incarcerated_10)

# 2010
# CO	Crowley County	45.6294007		
# TX	Concho County	38.0719354		
# LA	West Feliciana Parish	33.0944000		
# CO	Bent County	32.2818895		
# TX	Garza County	30.8775731	

mean(incarcerated_pop$percent_incarcerated_10) # = 1.458736

incarcerated_pop %>%
  arrange(desc(percent_incarcerated_00)) %>%
  select(state, county, percent_incarcerated_00)

# 2000
# CO	Crowley County	35.4295034		
# LA	West Feliciana Parish	33.2340679		
# TX	Concho County	32.7534039		
# FL	Union County	30.2112781		
# IL	Brown County	27.5107914

mean(incarcerated_pop$percent_incarcerated_00) # = 1.19643

# Findings:
# Average percent of incarcerated adult population has gone down nationwide since 2000. Crowley County and West Feliciana Parish are on every top five list.

```

### Create State Dataframes & Calculate Incarceration Rates

```{r}

# dataframes for each state w/ total pop, total incarcerated, state incarceration rate, percent of state incarcerated (2000-2020)

massachusetts <- subset(incarcerated_pop, state == "MA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

connecticut <- subset(incarcerated_pop, state == "CT") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

west_virginia <- subset(incarcerated_pop, state == "WV") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

vermont <- subset(incarcerated_pop, state == "VT") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

rhode_island <- subset(incarcerated_pop, state == "RI") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

washington_dc <- subset(incarcerated_pop, state == "DC") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

pennsylvania <- subset(incarcerated_pop, state == "PA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

arkansas <- subset(incarcerated_pop, state == "AR") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

oklahoma <- subset(incarcerated_pop, state == "OK") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

nebraska <- subset(incarcerated_pop, state == "NE") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

idaho <- subset(incarcerated_pop, state == "ID") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

missouri <- subset(incarcerated_pop, state == "MO") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

louisiana <- subset(incarcerated_pop, state == "LA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

utah <- subset(incarcerated_pop, state == "UT") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

arizona <- subset(incarcerated_pop, state == "AZ") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

south_carolina <- subset(incarcerated_pop, state == "SC") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

georgia <- subset(incarcerated_pop, state == "GA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

tennessee <- subset(incarcerated_pop, state == "TN") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

new_york <- subset(incarcerated_pop, state == "NY") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

mississippi <- subset(incarcerated_pop, state == "MS") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

new_mexico <- subset(incarcerated_pop, state == "NM") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

illinois <- subset(incarcerated_pop, state == "IL") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

michigan <- subset(incarcerated_pop, state == "MI") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

maine <- subset(incarcerated_pop, state == "ME") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

indiana <- subset(incarcerated_pop, state == "IN") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

north_carolina <- subset(incarcerated_pop, state == "NC") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

delaware <- subset(incarcerated_pop, state == "DE") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

texas <- subset(incarcerated_pop, state == "TX") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

alabama <- subset(incarcerated_pop, state == "AL") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

california <- subset(incarcerated_pop, state == "CA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

virginia <- subset(incarcerated_pop, state == "VA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

south_dakota <- subset(incarcerated_pop, state == "SD") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

montana <- subset(incarcerated_pop, state == "MT") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

new_jersey <- subset(incarcerated_pop, state == "NJ") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

hawaii <- subset(incarcerated_pop, state == "HI") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

ohio <- subset(incarcerated_pop, state == "OH") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

minnesota <- subset(incarcerated_pop, state == "MN") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

oregon <- subset(incarcerated_pop, state == "OR") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

north_dakota <- subset(incarcerated_pop, state == "ND") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

kentucky <- subset(incarcerated_pop, state == "KY") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

new_hampshire <- subset(incarcerated_pop, state == "NH") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

colorado <- subset(incarcerated_pop, state == "CO") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

washington <- subset(incarcerated_pop, state == "WA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

alaska <- subset(incarcerated_pop, state == "AK") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

wyoming <- subset(incarcerated_pop, state == "WY") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

iowa <- subset(incarcerated_pop, state == "IA") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

wisconsin <- subset(incarcerated_pop, state == "WI") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

florida <- subset(incarcerated_pop, state == "FL") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

kansas <- subset(incarcerated_pop, state == "KS") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

nevada <- subset(incarcerated_pop, state == "NV") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

maryland <- subset(incarcerated_pop, state == "MD") %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>% 
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000) %>%
  mutate(total_pct_incarcerated_20 = total_incarcerated_20/total_state_pop_20*100) %>% # 2010 below
  mutate(total_state_pop_10 = sum(total_population_10)) %>%
  mutate(total_incarcerated_10 = sum(incarcerated_10)) %>%
  mutate(state_incarceration_rate_10 = total_incarcerated_10/total_state_pop_10*100000) %>%
  mutate(total_pct_incarcerated_10 = total_incarcerated_10/total_state_pop_10*100) %>% # 2000 below
  mutate(total_state_pop_00 = sum(total_population_00)) %>%
  mutate(total_incarcerated_00 = sum(incarcerated_00)) %>%
  mutate(state_incarceration_rate_00 = total_incarcerated_10/total_state_pop_00*100000) %>%
  mutate(total_pct_incarcerated_00 = total_incarcerated_10/total_state_pop_00*100)

```

### 1. Which states had the highest incarceration rate per 100,000 people over time (2000-2020)? How does it compare to population growth?

```{r}

install.packages("usmap")
library(usmap)

incarceration_rates <- read_csv("data/2020-incarceration-rates.csv")

# top 10 highest incarceration rates in 2020
incarceration_rates %>%
  arrange(desc(incarceration_rate)) %>%
  head(10)

# Louisiana	-	1100.1220	
# West Virginia	- 1096.5500	
# Mississippi	- 1038.2340	
# Oklahoma - 971.2445	
# Arkansas - 	899.1793	
# Arizona	- 897.0703	
# Georgia	- 858.2225	
# Texas	- 853.5244	
# Kentucky - 851.0296	
# Alabama - 791.1384	

# top 10 lowest incarceration rates in 2020
incarceration_rates %>%
  arrange(incarceration_rate) %>%
  head(10)

# Vermont	- 189.5574	
# Maine	- 246.6310	
# Massachusetts	- 255.6076	
# Hawaii - 257.8214	
# Rhode Island - 280.8510	
# Minnesota	- 292.1584	
# New York - 310.4115	
# New Hampshire	-	319.0495	
# Utah - 326.4442	
# North Dakota	- 329.9987	

# Map
plot_usmap(data = incarceration_rates, values = "incarceration_rate", color = "black") + 
  labs(title = "Incarcerations Rates Across the United States in 2020",
       subtitle = "Louisiana has the highest incarceration rate at 1100 per 100,000 people.") +
  scale_fill_continuous(
    low = "white", high = "dark green", name = "Incarceration Rate", label = scales::comma) + theme(legend.position = "right")

```

### 2. Which U.S. region has the highest incarceration rates?

```{r}

# Northeast

plot_usmap(include = .northeast_region, data = incarceration_rates, values = "incarceration_rate", color = "black") + theme(legend.position = "right")

# Midwest

plot_usmap(include = .midwest_region, data = incarceration_rates, values = "incarceration_rate", color = "black") + theme(legend.position = "right")

# West

plot_usmap(include = .west_region, data = incarceration_rates, values = "incarceration_rate", color = "black") + theme(legend.position = "right")

# South 

plot_usmap(include = .south_region, data = incarceration_rates, values = "incarceration_rate", color = "black") + theme(legend.position = "right")

```





### 3. Which were the top 10 U.S. counties that saw the largest percent increase in prison population since 2010? Alternatively, which were the top 10 U.S. counties that saw either a decrease or no change in prison population since 2010? 

```{r}

# net change (direct subtraction over years)

percent_counties_00 <- incarcerated_pop %>%
  select(fips, percent_incarcerated_00)

percent_counties_10 <- incarcerated_pop %>%
  select(fips, percent_incarcerated_10)

percent_counties_20 <- incarcerated_pop %>%
  select(fips, percent_incarcerated_20)
  
# 2000

incarcerated_pop %>%
  arrange(desc(percent_incarcerated_00)) %>%
  select(state, county, percent_incarcerated_00) %>%
  head(10)

# county level map
plot_usmap(regions = "counties", data = percent_counties_00, values = "percent_incarcerated_00", color = "black") +
  labs(title = "Title",
       subtitle = "Subtitle") +
  scale_fill_continuous(low = "#c9d9c9", high = "#236d32", name = "Legend", label = scales::comma) + 
  theme(legend.position = "right")

# 2010

incarcerated_pop %>%
  arrange(desc(percent_incarcerated_10)) %>%
  select(state, county, percent_incarcerated_10) %>%
  head(10)

# county level map
plot_usmap(regions = "counties", data = percent_counties_10, values = "percent_incarcerated_10", color = "black") +
  labs(title = "Title",
       subtitle = "Subtitle") +
  scale_fill_continuous(low = "#c9d9c9", high = "#236d32", name = "Legend", label = scales::comma) + 
  theme(legend.position = "right")


# 2020

incarcerated_pop %>%
  arrange(desc(percent_incarcerated_20)) %>%
  select(state, county, percent_incarcerated_20) %>%
  head(10)

# county level map
plot_usmap(regions = "counties", data = percent_counties_20, values = "percent_incarcerated_20", color = "black") +
  labs(title = "Title",
       subtitle = "Subtitle") +
  scale_fill_continuous(low = "#c9d9c9", high = "#236d32", name = "Legend", label = scales::comma) + 
  theme(legend.position = "right")

```

### 4. For each U.S. state, roughly how much (%) of the states adult population was incarcerated/being held in a correctional facility in 2020? Compare it to figures from 2010 and 2000.

```{r}

```

### 5. Using ACS data, find out which racial or ethnic groups make up the majority in the top 10 U.S. counties that saw the largest increase in prison population since 2010?

```{r}

```


### Example County Level Maps (Percent Incarcerated)

```{r}

# Maryland

md_geo <- get_acs(geography = "county",
      variables = c(population = "B01001_001"),
      state = "MD",
      geometry = TRUE)

md_geo <- md_geo %>% 
  select(-variable, -estimate, -moe) %>%
  rename(geoid = GEOID) %>%
  rename(county = NAME) %>%
  mutate(county = str_remove_all(county,", Maryland"))

# total state pop + state incarceration rate 2020

maryland <- maryland %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>%
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000)

# join dataframes

maryland <- md_geo %>%
  left_join(maryland, by="county")

# map prep

counties <- counties()

md_counties <- counties %>%
  filter(STATEFP == "24")

# map

ggplot() + 
  geom_sf(data=maryland, aes(fill=percent_incarcerated_20)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma")


# Virginia

va_geo <- get_acs(geography = "county",
      variables = c(population = "B01001_001"),
      state = "VA",
      geometry = TRUE)

va_geo <- va_geo %>% 
  select(-variable, -estimate, -moe) %>%
  rename(geoid = GEOID) %>%
  rename(county = NAME) %>%
  mutate(county = str_remove_all(county,", Virginia"))

virginia <- virginia %>%
  mutate(total_state_pop_20 = sum(total_population_20)) %>%
  mutate(total_incarcerated_20 = sum(incarcerated_20)) %>%
  mutate(state_incarceration_rate_20 = total_incarcerated_20/total_state_pop_20*100000)

# join dataframes

virginia <- va_geo %>%
  left_join(virginia, by="county")

counties <- counties()

va_counties <- counties %>%
  filter(STATEFP == "51")


ggplot() + 
  geom_sf(data=virginia, aes(fill=percent_incarcerated_20)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma")

```
