---
title: "Reverse Engineering Project"
author: "Amanda Hernandez, Greg Morton"
date: "3/5/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban (https://www.wral.com/in-session-campaign-donations/17716104/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(tidyr)
library(lubridate)
library(dplyr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
nc_session_donations <- read.csv("data/nc_session_donations.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

# fix amount column
nc_session_donations <- nc_session_donations %>% 
  mutate(amount=as.numeric(parse_number(amount)))

# fix date column
nc_session_donations <- nc_session_donations %>%
  mutate(date = mdy(date)) 

# check data
View(nc_session_donations)

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Senate President Pro Tem Phil Berger's campaign alone raised nearly $222,000 during session. Twenty-three people gave him the maximum allowed: $5,200.
* **Analysis summary**: Our analysis found that Senate President Pro Tem Phil Berger's campaign raised $221,825 during the legislative session, confirming the quote "nearly $222,000" in the story. We were also able to confirm that 23 people donated exactly $5,200, which is the maximum amount allowed, to Berger's campaign.

```{r}
# Put code to reverse engineer sentence here

nc_session_donations %>%
  filter(committee_name == "PHILIP E BERGER COMM") %>%
  filter(date > "2018-05-16" & date < "2018-06-30") %>%
  group_by(committee_name) %>%
  summarize(total_amount = sum(amount)) # confirms Berger received nearly Senate President Pro Tem Phil Berger ($221,825 exactly)

berger_max_donors <- nc_session_donations %>%
  filter(committee_name == "PHILIP E BERGER COMM") %>%
  filter(date > "2018-05-16" & date < "2018-06-30") %>% 
  filter(amount == 5200) # confirms 23 people donated max amount to Berger

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: There was a major bump during the first week of session this year for the state parties, but that's because three large donors – LinkedIn co-founder Reid Hoffman, Durham investor Greg Lindberg and Lindberg's Eli Research company – donated $1.3 million that week to the state Republican and Democrat parties.
* **Analysis summary**: Our analysis confirmed that during the first week of the legislative session, Reid Hoffman, Greg Lindberg, and Lindberg's company Eli Research LLC were the top three donors. However, when adding up their respective donation amounts, the result is $1.25 million, not $1.3 million.

```{r}
# Put code to reverse engineer sentence here

first_wk_top_donors <- nc_session_donations %>%
  filter(date >= "2018-05-16" & date <= "2018-05-23") %>%
  group_by(name.clean) %>%
  summarise(total_donation_amount = sum(amount)) %>%
  arrange(desc(total_donation_amount)) # confirms Hoffman, Lindberg, and Eli Research LLC as top donors during first week

500000 + 500000 + 250000 # = 1,250,000 (~$1.3 mil donations between Hoffman, Lindberg, and Eli Research LLC)

```

### Sentence 3

* **Sentence text**: Beer distributors also gave during session. Mark Craig, the president of distributor R.H. Barringer, put more than $30,000 into state legislative campaigns during the six-week period, including the maximum $5,200 to Berger. He was the top individual donor to legislative campaigns during session.
* **Analysis summary**: Our analysis confirmed that Mark Craig donated the maximum amount allowed to Berger and donated over $30,000 total during the legislative session. We also confirmed that Craig was the top individual donor to legislative campaigns during the session, donating a total of $25,000. 

```{r}
# Put code to reverse engineer sentence here

mark_craig_donations <- nc_session_donations %>%
  filter(name.clean == "MARK R CRAIG") %>%
  filter(date >= "2018-05-16" & date <= "2018-06-30") %>%
  arrange(desc(amount)) # confirms max donation to Berger

sum(mark_craig_donations$amount) # confirms donation total over $30,000

# state party committees (committee_name) -> NC REC - STATE, NC REC BUILDING FUND, NORTH CAROLINA SENATE MAJORITY FUND, NC HOUSE REPUBLICAN CAMPAIGN COMMITTEE, NC REPUBLICAN COUNCIL OF STATE COMMITTEE, NC DEC - STATE, NORTH CAROLINA DEMOCRATIC PARTY - BUILDING

individual_top_donors <- nc_session_donations %>% 
  subset(committee_name!="NC DEC - STATE" & committee_name!="NC REC - STATE" & committee_name!="NC REC BUILDING FUND" & committee_name!="NORTH CAROLINA DEMOCRATIC PARTY - BUILDING" & committee_name!="NC REPUBLICAN COUNCIL OF STATE COMMITTEE" & committee_name!="NC HOUSE REPUBLICAN CAMPAIGN COMMITTEE" & committee_name!= "NORTH CAROLINA SENATE MAJORITY FUND") %>%
  filter(date > "2018-05-16" & date < "2018-06-30") %>%
  filter(transaction_type == "Individual") %>%
  group_by(name.clean) %>%
  summarise(total_donation_amount = sum(amount)) %>%
  arrange(desc(total_donation_amount)) # confirms Craig as individual top donor

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: R.H. Barringer's board chair, Jasie Barringer, donated another $8,200 in-session, and Teresa Craig kicked in another $5,200 to Berger. The North Carolina Beer & Wine Wholesalers Association PAC gave $18,400 during the week before session and on opening day.
* **Analysis summary**: Our analysis confirmed the donations amounts made by Jasie Barringer, Teresa Craig, and the North Carolina Beer & Wine Wholesalers Association PAC.

```{r}
# Put code to reverse engineer sentence here

# Jasie Barringer -> $8,200
nc_session_donations %>%
  filter(name.clean == "JASIE S BARRINGER") %>%
  filter(date >= "2018-05-16" & date <= "2018-06-30") %>%
  group_by(name.clean) %>%
  summarize(total_amount = sum(amount))

# Teresa Craig -> $5,200
nc_session_donations %>%
  filter(name.clean == "TERESA S CRAIG") %>%
  filter(date >= "2018-05-16" & date <= "2018-06-30") %>%
  group_by(name.clean) %>%
  summarize(total_amount = sum(amount))

# North Carolina Beer & Wine Wholesalers Association PAC -> $18,400
nc_session_donations %>%
  filter(name.clean == "NC BEER & WINE WHOLESALERS ASSOC PAC") %>%
  filter(date >= "2018-05-09" & date <= "2018-05-16") %>%
  group_by(name.clean) %>%
  summarize(total_amount = sum(amount))

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: The North Carolina Pork Council gave $10,000 during session to the state Republican Party and $5,000 right at the session's start to the Democrats.
* **Analysis summary**: Our analysis confirmed that the North Carolina Pork Council PAC donated $10k to the Republicans during the session and $5k to the Democratic Party at the beginning of the session.

```{r}
# Put code to reverse engineer sentence here

# republican party -> NC REC - STATE, democratic party -> NC DEC - STATE

nc_session_donations %>%
  filter(name.clean == "NC PORK COUNCIL PAC") %>%
  filter(committee_name == "NC REC - STATE") %>%
  filter(date >= "2018-05-16" & date <= "2018-06-30") %>%
  group_by(committee_name) %>%
  summarize(total_amount = sum(amount)) # confirms 10k donated to Republican Party during session

nc_session_donations %>%
  filter(name.clean == "NC PORK COUNCIL PAC") %>%
  filter(committee_name == "NC DEC - STATE") %>%
  filter(date >= "2018-05-15" & date <= "2018-05-16") %>%
  group_by(committee_name) %>%
  summarize(total_amount = sum(amount)) # confirms 5k donated to Democratic Party at the beginning of the session

# Display results of code below this codeblock

```

-30-