---
title: "lab_04"
author: "Amanda Hernandez"
date: "2/17/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load Data

Let's keep working with Maryland loan applications contained in the PPP data.

**Task** Read in the lab_04.rds file and assign it to the variable `maryland_ppp`.

```{r}
# add code here to accomplish the task

maryland_ppp <- readRDS("lab_04.rds")

```

To answer the questions below, you'll need to join both the NAICS code data and the ZCTA demographic data from the pre-lab. There's a particular set of steps you'll need to perform to make that work properly, so take your time and check your work.

First, let's load the NAICS code CSV file we have in the data folder:

**Task** Read in the naics_codes.csv file and assign it to the variable `naics_codes`.

```{r}
# add code here to accomplish the task

naics_codes <- read_csv('data/naics_codes.csv')

```

Next join the two and then create a 5-character `zip5` column that we can use to join to the ZCTA data:

**Task** Write code to join the maryland_ppp and naics_codes data, assigning the result to a new variable. Then add a new `zip5` column.

```{r}
# add code here to accomplish the task

md_ppp_naics <- maryland_ppp %>%
  left_join(naics_codes)

md_ppp_naics <- md_ppp_naics %>%
  mutate(zip5 = str_sub(zip, start=1L, end=5L))

```

Let's add the Zip Code Tabulation Area dataset. You'll need to make the `ZCTA5N` column a character column, not a numeric one.

**Task** Read in the Zip Code Tabulation Area dataset in the data folder and change the datatype of the `ZCTA5N` column to character. IMPORTANT: you can see what the column names represent by checking the data documentation, which can be found here: https://geodata.md.gov/imap/rest/services/Demographics/MD_CensusData/FeatureServer/1. You will need to refer to this to answer the questions below.

```{r}
# add code here to accomplish the task

md_zcta <- read_csv("data/maryland_zcta.csv")

md_zcta <- maryland_zcta %>% 
  mutate(across(ZCTA5N, as.character))

```

Now we can join the PPP data to the ZCTA data:

**Task** Join the Maryland PPP data (with NAICS already joined) with the ZCTA data, specifying the join column in both dataframes.

```{r}
# add code here to accomplish the task

md_ppp_naics_demographics <- md_ppp_naics %>%
  left_join(md_zcta, by=c("zip5"="ZCTA5N"))

```

## Answer questions

**Q1** What are the top 5 industries (using the NAICS title) with the most PPP applications in Maryland?
**A1** The top five industries include: 1. taxis service, 2. beauty salons, 3. full-service restaurants, 4. offices of real estate agents and brokers, and 5. all other personal services.

```{r}
# Write the code to produce your answer here

md_ppp_naics_demographics %>%
  group_by(title) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

```

**Q2** What are the top industries (using the NAICS title) with the most PPP applications in Maryland where the loan amount is more than $150,000?
**A2** The top five industries include: 1. full-service restaurants, 2. offices of physicians (except mental health specialists), 3. plumbing, heating, and air-conditioning contractors, 4. offices of dentists, and 5. commercial and institutional building construction.

```{r}
# Write the code to produce your answer here

md_ppp_naics_demographics %>%
  filter(amount > 150000) %>%
  group_by(title) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

```

**Q3** Which Maryland zip code with at least 20 percent non-Hispanic Asian population had the largest sum of PPP loans? Also show the number (count) of loans.
**A3** The zip code with the largest total amount of loans was 20850.

```{r}
# Write the code to produce your answer here

md_ppp_naics_demographics %>%
  filter(PNHA >= 20) %>%
  group_by(zip5) %>%
  summarise(count=n(), total_amount = sum(amount)) %>%
  arrange(desc(total_amount))

```

**Q4** Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian", and what is the percentage of non-Hispanic Asian population in that zip code? Also show the number (count) of loans.
**A4** The zip code is 20850 and the percentage of the non-Hispanic Asian population is 23.5 percent. The number of loans is 240.

```{r}
# Write the code to produce your answer here

md_ppp_naics_demographics %>%
  filter(race == "Asian") %>%
  group_by(zip5, PNHA) %>%
  summarise(count=n(), total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
  
```

**Q5** What might explain why 20850 has a significant Asian population (23.5%) and only 240 of the 3,186 loans (7.5%) are to Asian applicants?
**A5** Based off of past discussions about PPP race data, we know that some applicants chose not to provide their race or put "N/A" under the race field. There's also some discrepancies with how the race question on the application should have been filled out since it could refer to the race of the owner, CEO, etc. or the race of the person who filled the application out, which may not also be the owner. 

**Q6** Check to see if any Maryland zip codes that have loans where the PPP data lists the applicant as being Asian BUT are located in a zip code that does not appear to have Asian population according to the ZCTA data. How many loans do they have and what is the total amount? Take a look at these zip codes - what could explain this?
**A6** The zip codes include 21031 (total amount: $149,450), 21240 (total amount: $296,076), and 20701 (total amount: $10,400. The first two zip codes had two loans while the third had only one loan. One possibility that could explain this are errors in data entry where an applicant put down the incorrect race. It's also possible that the race fields were changed by mistake during the database creation process. Another possibility is that the applicant filled out the race field based on the race of the owner/CEO or their own race if they're not the same person.

```{r}
# Write the code to produce your answer here
  
md_ppp_naics_demographics %>%
  filter(PNHA == 0 & race == "Asian") %>%
  group_by(zip5, PNHA) %>%
  summarise(count = n(), total_amount = sum(amount)) %>%
  arrange(desc(count))

```

**Q7** How many industries have exactly one loan in the data (hint: you can filter _after_ using `summarise` on any totals you create). From a newsworthiness standpoint, which of these is most worthy of further exploration, and why?
**A7** 44 industries have exactly one loan. I would want to learn more about the ammunition manufacturing industry because I feel like it would be the most interesting to readers. I also know that at the beginning of the pandemic a lot of people were "doomsday prepping," so I'm curious whether that's related to the ammunition industry at all.

```{r}
# Write the code to produce your answer here

md_ppp_naics_demographics %>%
  group_by(title) %>%
  summarise(count = n()) %>%
  filter(count == 1) %>%
  arrange(desc(count))

```
